---
title: "Journal (reproducible report)"
author: "Jader Santana"
date: "2020-11-23"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    toc_depth: 3
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```
# Challenge 01_intro_r Introduction to R, RStudio IDE & Github
no specific challenge for this unit

# Challenge 02_intro_tv Intro to tidyverse

Last compiled: `r Sys.time()`

## 1.0 Load libraries ----
```{r}
library(tidyverse)
#  library(tibble)    --> is a modern re-imagining of the data frame
#  library(readr)     --> provides a fast and friendly way to read rectangular data like csv
#  library(dplyr)     --> provides a grammar of data manipulation
#  library(magrittr)  --> offers a set of operators which make your code more readable (pipe operator)
#  library(tidyr)     --> provides a set of functions that help you get to tidy data
#  library(stringr)   --> provides a cohesive set of functions designed to make working with strings as easy as possible
#  library(ggplot2)   --> graphics


# Excel Files
library(readxl)
```

## 2.0 Importing Files ----
```{r}
# A good convention is to use the file name and suffix it with tbl for the data structure tibble
bikes_tbl      <- read_excel(path = "00_data/01_bike_sales/01_raw_data/bikes.xlsx")
orderlines_tbl <- read_excel("00_data/01_bike_sales/01_raw_data/orderlines.xlsx")

# Not necessary for this analysis, but for the sake of completeness
bikeshops_tbl  <- read_excel("00_data/01_bike_sales/01_raw_data/bikeshops.xlsx")
```
## 3.0 Examining Data ----
```{r}
bikeshops_tbl
glimpse(bikeshops_tbl)
```
## 4.0 Joining Data ----
```{r}
left_join(orderlines_tbl, bikes_tbl, by = c("product.id" = "bike.id"))
bike_orderlines_joined_tbl <- orderlines_tbl %>%
  left_join(bikes_tbl, by = c("product.id" = "bike.id")) %>%
  left_join(bikeshops_tbl, by = c("customer.id" = "bikeshop.id"))
bike_orderlines_joined_tbl %>% glimpse()
```
## 5.0 Wrangling Data ----
```{r}
# All actions are chained with the pipe already. You can perform each step separately and use glimpse() or View() to validate your code. Store the result in a variable at the end of the steps.
bike_orderlines_wrangled_tbl <- bike_orderlines_joined_tbl %>%
  # 5.1 Separate category name
  separate(col    = location,
           into   = c("city", "state"),
           sep    = ", ") %>%
  
  # 5.2 Add the total price (price * quantity) 
  # Add a column to a tibble that uses a formula-style calculation of other columns
  mutate(total.price = price * quantity) %>%
  
  # 5.3 Optional: Reorganize. Using select to grab or remove unnecessary columns
  # 5.3.1 by exact column name
  select(-...1, -gender) %>%
  
  # 5.3.2 by a pattern
  # You can use the select_helpers to define patterns. 
  # Type ?ends_with and click on Select helpers in the documentation
  select(-ends_with(".id")) %>%
  
  # 5.3.3 Actually we need the column "order.id". Let's bind it back to the data
  bind_cols(bike_orderlines_joined_tbl %>% select(order.id)) %>% 
  
  # 5.3.4 You can reorder the data by selecting the columns in your desired order.
  # You can use select_helpers like contains() or everything()
  select(order.id, contains("order"), contains("model"), contains("category"),
         price, quantity, total.price,
         everything()) %>%
  
  # 5.4 Rename columns because we actually wanted underscores instead of the dots
  # (one at the time vs. multiple at once)
  rename(bikeshop = name) %>%
  set_names(names(.) %>% str_replace_all("\\.", "_"))
```
## 6.0 Business Insights ----

### 6.1 Sales by Location ----
```{r}
library(lubridate)
# Step 1 - Manipulate
sales_by_location_tbl <- bike_orderlines_wrangled_tbl %>%
  
  # Select columns
  select(state, total_price) %>%
  
  # Add year column
  #no need mutate(year = year(order_date)) %>%
  
  # Grouping by state and summarizing sales
  group_by(state) %>% 
  summarize(sales = sum(total_price)) %>%
  
  # Optional: Add a column that turns the numbers into a currency format 
  # (makes it in the plot optically more appealing)
  # mutate(sales_text = scales::dollar(sales)) <- Works for dollar values
  mutate(sales_text = scales::dollar(sales, big.mark = ".", 
                                     decimal.mark = ",", 
                                     prefix = "", 
                                     suffix = " €"))

sales_by_location_tbl

# Step 2 - Visualize
sales_by_location_tbl %>%
  
  # Setup canvas with the columns location (x-axis) and sales (y-axis)
  ggplot(aes(x = state, y = sales)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  # Geometries
  geom_col(fill = "#2DC6D6") + # Use geom_col for a bar plot
  geom_label(aes(label = sales_text)) + # Adding labels to the bars
  geom_smooth(method = "lm", se = FALSE) + # Adding a trendline
  
  # Formatting
  # scale_y_continuous(labels = scales::dollar) + # Change the y-axis. 
  # Again, we have to adjust it for euro values
  scale_y_continuous(labels = scales::dollar_format(big.mark = ".", 
                                                    decimal.mark = ",", 
                                                    prefix = "", 
                                                    suffix = " €")) +
  labs(
    title    = "Revenue by Location(state)",
    subtitle = "Not an upward Trend",
    x = "", # Override defaults for x and y
    y = "Revenue"
  )

```

### 6.2 Sales by Year and Location ----

```{r}
# Step 1 - Manipulate
sales_by_loc_year_tbl <- bike_orderlines_wrangled_tbl %>%
  
  # Select columns and add a year
  select(order_date, total_price, state) %>%
  mutate(year = year(order_date)) %>%
  
  # Group by and summarize year and main catgegory
  group_by(year, state) %>%
  summarise(sales = sum(total_price)) %>%
  ungroup() %>%
  
  # Format $ Text
  mutate(sales_text = scales::dollar(sales, big.mark = ".", 
                                     decimal.mark = ",", 
                                     prefix = "", 
                                     suffix = " €"))

sales_by_loc_year_tbl
```
```{r plot, fig.width=13, fig.height=7}
# Step 2 - Visualize
sales_by_loc_year_tbl %>%
  
  # Set up x, y, fill
  ggplot(aes(x = year, y = sales, fill = state)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  # Geometries
  geom_col() + # Run up to here to get a stacked bar plot
  
  # Facet
  facet_wrap(~ state) +
  
  # Formatting
  scale_y_continuous(labels = scales::dollar_format(big.mark = ".", 
                                                    decimal.mark = ",", 
                                                    prefix = "", 
                                                    suffix = " €")) +
  labs(
    title = "Revenue by year and location",
    subtitle = "Each location has an upward trend",
    fill = "Main Location" # Changes the legend name
  )
```

# Challenge 03_Data Aquisition

Last compiled: `r Sys.time()`

## 1.0 Load libraries ----
test 2

# Challenge 04_Data Wrangling
Answer the following questions with that data:

## 1.0 Patent Dominance ----
1. Patent Dominance: What US company / corporation has the most patents? List the 10 US companies with the most assigned/granted patents.
```{r 1.0 Patent Dominance}
#install.packages("vroom")
#install.packages("tictoc")

# Tidyverse
library(tidyverse)
library(vroom)

# Data Table
library(data.table)

# Counter
library(tictoc)



col_types_assignee_tbl <- list(
  id = col_character(),
  type = col_double(),
  name_first = col_character(),
  name_last = col_character(),
  organization = col_character()
)

assignee_tbl <- vroom(
            file       = "00_data/04_challenge/assignee.tsv", 
            delim      = "\t", 
            col_types  = col_types_assignee_tbl,
            na         = c("", "NA", "NULL")
        )

#col_types_patent_tbl <- list(
 # id = col_character(),
  #type = col_character(),
  #number = col_character(),
  #country = col_character(),
 # date = col_date("%Y-%m-%d"),
#  abstract = col_character(),
  #title = col_character(),
 # kind = col_character(),
#  num_claims = col_double(),
  #filename = col_character(),
 # withdrawn = col_double()
#)

#patent_tbl <- vroom(
#            file       = "00_data/04_challenge/patent.tsv", 
#            delim      = "\t", 
#            col_types  = col_types_patent_tbl,
#            na         = c("", "NA", "NULL")
#        )

col_types_patent_assignee_tbl <- list(
  patent_id = col_character(),
  assignee_id = col_character(),
  location_id = col_character()
)

patent_assignee_tbl <- vroom(
            file       = "00_data/04_challenge/patent_assignee.tsv", 
            delim      = "\t", 
            col_types  = col_types_patent_assignee_tbl,
            na         = c("", "NA", "NULL")
        )

#col_types_uspc_tbl <- list(
 # uuid = col_character(),
  #patent_id = col_character(),
  #mainclass_id = col_character(),
  #subclass_id = col_character(),
  #sequence = col_double()
#)

#uspc_tbl <- vroom(
 #           file       = "00_data/04_challenge/uspc.tsv", 
  #          delim      = "\t", 
   #         col_types  = col_types_uspc_tbl,
    #        na         = c("", "NA", "NULL")
     #   )

assignee_tbl %>% glimpse()
#patent_tbl %>% glimpse()
patent_assignee_tbl %>% glimpse()
#uspc_tbl %>% glimpse()

class(assignee_tbl)
#class(patent_tbl)
class(patent_assignee_tbl)
#class(uspc_tbl)

setDT(assignee_tbl)
#setDT(patent_tbl)
setDT(patent_assignee_tbl)
#setDT(uspc_tbl)

class(assignee_tbl)
#class(patent_tbl)
class(patent_assignee_tbl)
#class(uspc_tbl)

assignee_tbl %>% glimpse()
#patent_tbl %>% glimpse()
patent_assignee_tbl %>% glimpse()
#uspc_tbl %>% glimpse()

# 4.0 DATA WRANGLING ----
# 4.1 Joining / Merging Data ----

tic()
combined_assignee_patent_assignee_tbl <- merge(x = assignee_tbl, y = patent_assignee_tbl, 
                       by.x = "id", by.y = "assignee_id",
                       all.x = TRUE, 
                       all.y = TRUE)%>%
mutate(patents_unit=1)%>%
 group_by(organization) %>%
  summarize(number_of_patents = sum(patents_unit)) %>%
  ungroup()%>%
  arrange(number_of_patents)
toc()

combined_assignee_patent_assignee_tbl %>%
    select(organization, number_of_patents) %>%
    arrange(desc(number_of_patents)) %>%
    slice(1:11) %>%
    filter(organization!="")

```


## 2.0 Recent Patent Acitivity ----
2. Recent patent activity: What US company had the most patents granted in 2019? List the top 10 companies with the most new granted patents for 2019.

```{r}

```


## 3.0 Innovation in Tech ----
3. Innovation in Tech: What is the most innovative tech sector? For the top 10 companies (worldwide) with the most patents, what are the top 5 USPTO tech main classes?

```{r}
library(tidyverse)
```
