---
title: "Unit 7"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Theory Input

## 1. Modeling

```{r I. Resources, eval=FALSE, include=FALSE}
#install.packages("tidymodels")
#install.packages("broom.mixed")

library(tidyverse)
library(tidyquant)
library(broom)
library(umap)
library(dplyr)

library(tidymodels)  # for the parsnip package, along with the rest of tidymodels

# Helper packages
library(broom.mixed) # for converting bayesian models to tidy tibbles

# Data set
bike_data_tbl_import <- readRDS("00_data/01_bike_sales/02_wrangled_data/bike_orderlines.rds") 

bike_data_tbl <- bike_data_tbl_import %>% select(everything()) %>% filter(category_1 != "Gravel")

ggplot(bike_data_tbl,
       aes(x = price, 
           y = weight, 
           group = category_1, 
           col = category_1)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE) +
  scale_color_manual(values=c("#2dc6d6", "#d65a2d", "#d6af2d", "#8a2dd6","#641E16"))

```

```{r II. Build and fit a model,eval=FALSE, include=FALSE}

#install.packages("parnsnip")

library(tidymodels)

weight ~ price * category_1

linear_reg()

lm_mod <- linear_reg() %>% 
              set_engine("lm")
              
lm_mod

lm_fit <- lm_mod %>% 
            fit(weight ~ price * category_1, 
                data = bike_data_tbl)
tidy(lm_fit)

new_points <- expand.grid(price = 20000, 
                          category_1 = c("E-Bikes", "Hybrid / City", "Mountain", "Road"))
new_points

mean_pred <- predict(lm_fit, new_data = new_points)
mean_pred

conf_int_pred <- predict(lm_fit, 
                         new_data = new_points, 
                         type = "conf_int")
conf_int_pred

plot_data <- new_points %>% 
              bind_cols(mean_pred) %>% 
              bind_cols(conf_int_pred)
ggplot(plot_data, aes(x = category_1)) + 
  geom_point(aes(y = .pred)) + 
  geom_errorbar(aes(ymin = .pred_lower, 
                    ymax = .pred_upper),
                width = .2) + 
  labs(y = "Bike weight", x = "Category") 

#install.packages("rstanarm")

library(rstanarm)
# set the prior distribution
prior_dist <- rstanarm::student_t(df = 1)

set.seed(123)

# make the parsnip model
bayes_mod <- linear_reg() %>% 
              set_engine("stan",
              prior_intercept = prior_dist, 
              prior = prior_dist) 

# train the model

bayes_fit <-  bayes_mod %>% 
                fit(weight ~ price * category_1, 
                data = bike_data_tbl)

print(bayes_fit, digits = 5)

tidy(bayes_fit, conf.int = TRUE)

bayes_plot_data <- 
  new_points %>%
  bind_cols(predict(bayes_fit, new_data = new_points)) %>% 
  bind_cols(predict(bayes_fit, new_data = new_points, type = "conf_int"))

ggplot(bayes_plot_data, aes(x = category_1)) + 
  geom_point(aes(y = .pred)) + 
  geom_errorbar(aes(ymin = .pred_lower, ymax = .pred_upper), width = .2) + 
  labs(y = "Bike weight") + 
  ggtitle("Bayesian model with t(1) prior distribution")

## parsnip model object
## 
## Fit time:  1m 39.2s 
## stan_glm
##  family:       gaussian [identity]
##  formula:      weight ~ price * category_1
##  observations: 218
##  predictors:   8
## ------
##                                    Median   MAD_SD  
## (Intercept)                        15.24134  1.25450
## price                          0.00158  0.00032
## category_1Hybrid / City            -2.34716  1.53781
## category_1Mountain                 -1.52799  1.31226
## category_1Road                     -6.47532  1.31294
## price:category_1Hybrid / City -0.00310  0.00066
## price:category_1Mountain      -0.00180  0.00035
## price:category_1Road          -0.00180  0.00033
## 
## Auxiliary parameter(s):
##       Median  MAD_SD 
## sigma 1.63413 0.07824
## 
## ------
## * For help interpreting the printed output see ?print.stanreg
## * For info on the priors used see ?prior_summary.stanreg

```

## 2. Preprocessing

### I. The data

```{r I. The data, eval=FALSE, include=FALSE}

#install.packages("nycflights13")
#install.packages("skimr")

library(nycflights13)
library(skimr)

set.seed(123)

flight_data <- 
  flights %>% 
  mutate(
    # Convert the arrival delay to a factor
    arr_delay = ifelse(arr_delay >= 30, "late", "on_time"),
    arr_delay = factor(arr_delay),
    # We will use the date (not date-time) in the recipe below
    date = as.Date(time_hour)
  ) %>% 
  # Include the weather data
  inner_join(weather, by = c("origin", "time_hour")) %>% 
  # Only retain the specific columns we will use
  select(dep_time, flight, origin, dest, air_time, distance, 
         carrier, date, arr_delay, time_hour) %>% 
  # Exclude missing data
  na.omit() %>% 
  # For creating models, it is better to have qualitative columns
  # encoded as factors (instead of character strings)
  mutate_if(is.character, as.factor)

flight_data %>% 
  count(arr_delay) %>% 
  mutate(prop = n/sum(n))

glimpse(flight_data)

flight_data %>% 
  skimr::skim(dest, carrier)





```

### II. Data Splitting   
```{r eval=FALSE, include=FALSE}
## Data Splitting

#install.packages("rsample")
library(rsample)

# Fix the random numbers by setting the seed 
# This enables the analysis to be reproducible when random numbers are used 
set.seed(555)
# Put 3/4 of the data into the training set 
data_split <- initial_split(flight_data, prop = 3/4)

# Create data frames for the two sets:
train_data <- training(data_split)
test_data  <- testing(data_split)
```

### III. Create recipe and roles

```{r III. Create recipe and roles, eval=FALSE, include=FALSE}

library(recipes)

flights_rec <- 
  recipe(arr_delay ~ ., data = train_data) 

flights_rec <- 
  recipe(arr_delay ~ ., data = train_data) %>% 
  update_role(flight, time_hour, new_role = "ID") 

summary(flights_rec)

```

### IV. Create features

```{r IV. Create features, eval=FALSE, include=FALSE}


flight_data %>% 
  distinct(date) %>% 
  mutate(numeric_date = as.numeric(date)) 

flights_rec <- 
 recipe(arr_delay ~ ., data = train_data) %>% 
 update_role(flight, time_hour, new_role = "ID") %>% 
 step_date(date, features = c("dow", "month")) %>%               
 step_holiday(date, holidays = timeDate::listHolidays("US")) %>% 
 step_rm(date)

test_data %>% 
  distinct(dest) %>% 
  anti_join(train_data)

flights_rec <- 
  recipe(arr_delay ~ ., data = train_data) %>% 
  update_role(flight, time_hour, new_role = "ID") %>% 
  step_date(date, features = c("dow", "month")) %>% 
  step_holiday(date, holidays = timeDate::listHolidays("US")) %>% 
  step_rm(date) %>% 
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_zv(all_predictors())

```

### V. Fit a model with a recipe

```{r V. Fit a model with a recipe, eval=FALSE, include=FALSE}

lr_mod <- 
  logistic_reg() %>% 
  set_engine("glm")

library(workflows)
library(parsnip)

flights_wflow <- 
  workflow() %>% 
  add_model(lr_mod) %>% 
  add_recipe(flights_rec)
flights_wflow

flights_fit <- 
  flights_wflow %>% 
  fit(data = train_data)

flights_fit %>% 
  pull_workflow_fit() %>% 
  tidy()

predict(flights_fit, test_data)

flights_pred <- 
  predict(flights_fit, test_data, type = "prob") %>% 
  bind_cols(test_data %>% select(arr_delay, time_hour, flight)) 

# The data look like: 
flights_pred

flights_pred %>% 
  roc_curve(truth = arr_delay, .pred_late) %>% 
  autoplot()

flights_pred %>% 
  roc_auc(truth = arr_delay, .pred_late)


```

## 3. Evaluating

```{r eval=FALSE, include=FALSE}
library(tidymodels)

library(modeldata)

```
### I. The Data

```{r 3. I. The Data, eval=FALSE, include=FALSE}
data(cells, package = "modeldata")
cells

cells %>% 
  count(class) %>% 
  mutate(prop = n/sum(n))
```
### II. Data splitting
```{r II. Data splitting, eval=FALSE, include=FALSE}
set.seed(123)
cell_split <- initial_split(cells %>% select(-case), 
                            strata = class)
cell_train <- training(cell_split)
cell_test  <- testing(cell_split)

nrow(cell_train)
## 1515
nrow(cell_train)/nrow(cells)
## 0.7503715

# training set proportions by class
cell_train %>% 
  count(class) %>% 
  mutate(prop = n/sum(n))
## # A tibble: 2 x 3
##   class     n  prop
##   <fct> <int> <dbl>
## 1 PS      975 0.644
## 2 WS      540 0.356  

# test set proportions by class
cell_test %>% 
  count(class) %>% 
  mutate(prop = n/sum(n))
## # A tibble: 2 x 3
##   class     n  prop
##   <fct> <int> <dbl>
## 1 PS      325 0.645
## 2 WS      179 0.355
```
### III. Modeling

```{r III. Modeling, eval=FALSE, include=FALSE}

#install.packages("ranger")
library(ranger)

rf_mod <- 
  rand_forest(trees = 1000) %>%
  set_engine("ranger") %>% 
  set_mode("classification")

set.seed(234)
rf_fit <- 
  rf_mod %>% 
  fit(class ~ ., data = cell_train)
rf_fit

rf_training_pred <- 
  predict(rf_fit, cell_train) %>% 
  bind_cols(predict(rf_fit, cell_train, type = "prob")) %>% 
  # Add the true outcome data back in
  bind_cols(cell_train %>% 
              select(class))

rf_training_pred %>%                # training set predictions
  roc_auc(truth = class, .pred_PS)
## # A tibble: 1 x 3
##   .metric .estimator .estimate
##   <chr>   <chr>          <dbl>
## 1 roc_auc binary          1.00
rf_training_pred %>%                # training set predictions
  accuracy(truth = class, .pred_class)
## # A tibble: 1 x 3
##   .metric  .estimator .estimate
##   <chr>    <chr>          <dbl>
## 1 accuracy binary         0.993  

rf_testing_pred <- 
  predict(rf_fit, cell_test) %>% 
  bind_cols(predict(rf_fit, cell_test, type = "prob")) %>% 
  bind_cols(cell_test %>% select(class))

rf_testing_pred %>%                   # test set predictions
  roc_auc(truth = class, .pred_PS)
## # A tibble: 1 x 3
##   .metric .estimator .estimate
##   <chr>   <chr>          <dbl>
## 1 roc_auc binary         0.909

rf_testing_pred %>%                   # test set predictions
  accuracy(truth = class, .pred_class)
## # A tibble: 1 x 3
##   .metric  .estimator .estimate
##   <chr>    <chr>          <dbl>
## 1 accuracy binary         0.837
```
### IV. Estimating performance
```{r IV. Estimating performance}



```
### V. Resampling to the rescue
```{V. Resampling to the rescue}

```
### VI. Fit a model with resampling
```{r VI. Fit a model with resampling}

```

## 4. Tuning
```{r 4. Tuning}

```
### I. The cell image data, revisited
```{r I. The cell image data, revisited}

```
### II. Predicting image segmentation, but better
```{r II. Predicting image segmentation, but better}

```
### III. Tuning hyperparameters
```{r III. Tuning hyperparameters}

```
### IV. Model tuning with a grid
```{r IV. Model tuning with a grid}

```
### V. Finalizing our model
```{r V. Finalizing our model}

```
### VI. Exploring results
```{r VI. Exploring results}

```
### VII. The last fit
```{r VII. The last fit}

```


# Chapter 8 - Business Case

```{r eval=FALSE, include=FALSE}
# Standard
library(tidyverse)

# Modeling
library(parsnip)

# Preprocessing & Sampling
library(recipes)
library(rsample)

# Modeling Error Metrics
library(yardstick)

#install.packages("rpart.plot")

# Plotting Decision Trees
library(rpart.plot)


```
### 
```{r eval=FALSE, include=FALSE}
# Modeling ----------------------------------------------------------------
bike_orderlines_tbl <- readRDS("00_data/01_bike_sales/01_raw_data/bike_orderlines.rds")
glimpse(bike_orderlines_tbl)

model_sales_tbl <- bike_orderlines_tbl %>%
    select(total_price, model, category_2, frame_material) %>%
    
    group_by(model, category_2, frame_material) %>%
    summarise(total_sales = sum(total_price)) %>%
    filter(!(category_2 %in% c("Downhill", "Adventure", "E-Road", "Fat Bikes"))) %>%
    ungroup() %>%
    
    arrange(desc(total_sales))

model_sales_tbl %>%
    mutate(category_2 = as_factor(category_2) %>% 
               fct_reorder(total_sales, .fun = max) %>% 
               fct_rev()) %>%
    
    ggplot(aes(frame_material, total_sales)) +
    geom_violin() +
    geom_jitter(width = 0.1, alpha = 0.5, color = "#2c3e50") +
    #coord_flip() +
    facet_wrap(~ category_2) +
    scale_y_continuous(labels = scales::dollar_format(scale = 1e-6, suffix = "M", accuracy = 0.1)) +
    tidyquant::theme_tq() +
    labs(
        title = "Total Sales for Each Model",
        x = "Frame Material", y = "Revenue"
    )
```
### 
```{r eval=FALSE, include=FALSE}
bike_features_tbl <- readRDS("00_data/01_bike_sales/01_raw_data/bike_features_tbl.rds")
glimpse(bike_features_tbl)

bike_features_tbl <- bike_features_tbl %>% 
    select(model:url, `Rear Derailleur`, `Shift Lever`) %>% 
    mutate(
      `shimano dura-ace`        = `Rear Derailleur` %>% str_to_lower() %>% str_detect("shimano dura-ace ") %>% as.numeric(),
      `shimano ultegra`         = `Rear Derailleur` %>% str_to_lower() %>% str_detect("shimano ultegra ") %>% as.numeric(),
      `shimano 105`             = `Rear Derailleur` %>% str_to_lower() %>% str_detect("shimano 105 ") %>% as.numeric(),
      `shimano tiagra`          = `Rear Derailleur` %>% str_to_lower() %>% str_detect("shimano tiagra ") %>% as.numeric(),
      `Shimano sora`            = `Rear Derailleur` %>% str_to_lower() %>% str_detect("shimano sora") %>% as.numeric(),
      `shimano deore`           = `Rear Derailleur` %>% str_to_lower() %>% str_detect("shimano deore(?! xt)") %>% as.numeric(),
      `shimano slx`             = `Rear Derailleur` %>% str_to_lower() %>% str_detect("shimano slx") %>% as.numeric(),
      `shimano grx`             = `Rear Derailleur` %>% str_to_lower() %>% str_detect("shimano grx") %>% as.numeric(),
      `Shimano xt`              = `Rear Derailleur` %>% str_to_lower() %>% str_detect("shimano deore xt |shimano xt ") %>% as.numeric(),
      `Shimano xtr`             = `Rear Derailleur` %>% str_to_lower() %>% str_detect("shimano xtr") %>% as.numeric(),
      `Shimano saint`           = `Rear Derailleur` %>% str_to_lower() %>% str_detect("shimano saint") %>% as.numeric(),
      `SRAM red`                = `Rear Derailleur` %>% str_to_lower() %>% str_detect("sram red") %>% as.numeric(),
      `SRAM force`              = `Rear Derailleur` %>% str_to_lower() %>% str_detect("sram force") %>% as.numeric(),
      `SRAM rival`              = `Rear Derailleur` %>% str_to_lower() %>% str_detect("sram rival") %>% as.numeric(),
      `SRAM apex`               = `Rear Derailleur` %>% str_to_lower() %>% str_detect("sram apex") %>% as.numeric(),
      `SRAM xx1`                = `Rear Derailleur` %>% str_to_lower() %>% str_detect("sram xx1") %>% as.numeric(),
      `SRAM x01`                = `Rear Derailleur` %>% str_to_lower() %>% str_detect("sram x01|sram xo1") %>% as.numeric(),
      `SRAM gx`                 = `Rear Derailleur` %>% str_to_lower() %>% str_detect("sram gx") %>% as.numeric(),
      `SRAM nx`                 = `Rear Derailleur` %>% str_to_lower() %>% str_detect("sram nx") %>% as.numeric(),
      `SRAM sx`                 = `Rear Derailleur` %>% str_to_lower() %>% str_detect("sram sx") %>% as.numeric(),
      `SRAM sx`                 = `Rear Derailleur` %>% str_to_lower() %>% str_detect("sram sx") %>% as.numeric(),
      `Campagnolo potenza`      = `Rear Derailleur` %>% str_to_lower() %>% str_detect("campagnolo potenza") %>% as.numeric(),
      `Campagnolo super record` = `Rear Derailleur` %>% str_to_lower() %>% str_detect("campagnolo super record") %>% as.numeric(),
      `shimano nexus`           = `Shift Lever`     %>% str_to_lower() %>% str_detect("shimano nexus") %>% as.numeric(),
      `shimano alfine`          = `Shift Lever`     %>% str_to_lower() %>% str_detect("shimano alfine") %>% as.numeric()
    ) %>% 
  # Remove original columns  
  select(-c(`Rear Derailleur`, `Shift Lever`)) %>% 
  # Set all NAs to 0
  mutate_if(is.numeric, ~replace(., is.na(.), 0))

# 2.0 TRAINING & TEST SETS ----
bike_features_tbl <- bike_features_tbl %>% 
  
  mutate(id = row_number()) %>% 
  
  select(id, everything(), -url)

```
### 
```{r eval=FALSE, include=FALSE}

```


